---
title: '5'
author: "Jericho C. Gregory"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
hotel_bookings <- read_csv("hotel_bookings.csv",
                           col_types = cols(reservation_status_date = col_date(format = "%Y-%m-%d")))
summary(hotel_bookings)

str(hotel_bookings)
library(tidyverse)
library(lubridate)
# Clean data set
myHotel_bookings<-hotel_bookings%>%
  mutate(arrival_date=paste0(arrival_date_day_of_month,
                             arrival_date_month,
                             arrival_date_year)%>%
           dmy(),
         arrival_status_diff=arrival_date-reservation_status_date,
         reservation_status_date_month=month(reservation_status_date),
         reservation_status_date_year=year(reservation_status_date),
         arrival_date_month_num=month(arrival_date))%>% # clean dates
  mutate(hotel=as.factor(hotel),
         meal = as.factor(meal),
         country = as.factor(country),
         market_segment = as.factor(market_segment),
         distribution_channel = as.factor(distribution_channel),
         reserved_room_type = as.factor(reserved_room_type),
         assigned_room_type = as.factor(assigned_room_type),
         deposit_type = as.factor(deposit_type),
         agent = as.factor(agent),
         company = as.factor(company),
         reservation_status = as.factor(reservation_status),
         customer_type =as.factor(customer_type))%>% # convert character variables to factors
  mutate(reservation_date = arrival_date -ddays(lead_time),
         is_LMC = as.numeric(is_canceled==1 & arrival_status_diff<=30)) # engineer new variables
# Find LMCs by month
myHotel_LMCs<-myHotel_bookings%>%
  filter(is_LMC==1)%>% # select only LMCs
  group_by(hotel, reservation_status_date_year,reservation_status_date_month)%>%
  summarise(cancelations=sum(is_LMC))%>% # get total LMCs by month, year, and hotel
  arrange(hotel, reservation_status_date_year,reservation_status_date_month) # sort the data

print(myHotel_LMCs)

myHotel_bookings_completed_Jul17 <- myHotel_bookings%>%
  filter(reservation_status_date<ymd("20170701") | arrival_date<ymd("20170701"))
myHotel_bookings_outstanding_Jul17 <- myHotel_bookings%>%
  filter(reservation_status_date>=ymd("20170701") &
           arrival_date>=ymd("20170701") &
           reservation_date<ymd("20170701"))%>%
  select(-is_canceled,
         -reservation_status,
         -reservation_status_date,
         -reservation_status_date_month,
         -reservation_status_date_year,
         -arrival_status_diff,
         -assigned_room_type)
myHotel_bookings_completed_Aug17 <- myHotel_bookings%>%
  filter(reservation_status_date<ymd("20170801") | arrival_date<ymd("20170801"))
myHotel_bookings_outstanding_Aug17 <- myHotel_bookings%>%
  filter(reservation_status_date>=ymd("20170801") &
           arrival_date>=ymd("20170801") &
           reservation_date<ymd("20170801"))%>%
  select(-is_canceled,
         -reservation_status,
         -reservation_status_date,
         -reservation_status_date_month,
         -reservation_status_date_year,
         -arrival_status_diff,
         -assigned_room_type)

print(myHotel_LMCs)
```

```{r}
library(readr)
library(tidyverse)
library(lubridate)
library(forecast)
library(prophet)

july_train <- myHotel_bookings_completed_Jul17 
aug_train <- myHotel_bookings_completed_Aug17 
july_test <- myHotel_bookings_outstanding_Jul17
aug_test <- myHotel_bookings_outstanding_Aug17

modelTester <- function(x1, x2){
  return(sqrt((((x1-193)^2 + (x2-135)^2)) / 2))
}
```

```{r}
df_july <- myHotel_bookings_completed_Jul17 %>%
  filter(hotel == "Resort Hotel") %>%
  arrange(reservation_status_date_year, reservation_status_date_month) %>%
  group_by(reservation_status_date_year, reservation_status_date_month) %>%
  summarise(LMCs = sum(is_LMC)) %>%
  mutate(ds = paste(reservation_status_date_year, reservation_status_date_month, "01", sep = "-"),
         ds = as.Date(ds)) %>%
  select(ds, y = LMCs)

m_july <- prophet(df_july)

future_july <- make_future_dataframe(m_july, periods = 1, freq = "month")

forecast_july <- predict(m_july, future_july)

july_2017_forecast <- forecast_july$yhat[length(forecast_july$yhat)]

df_august <- myHotel_bookings_completed_Aug17 %>%
  filter(hotel == "Resort Hotel") %>%
  arrange(reservation_status_date_year, reservation_status_date_month) %>%
  group_by(reservation_status_date_year, reservation_status_date_month) %>%
  summarise(LMCs = sum(is_LMC)) %>%
  mutate(ds = paste(reservation_status_date_year, reservation_status_date_month, "01", sep = "-"),
         ds = as.Date(ds)) %>%
  select(ds, y = LMCs)

m_august <- prophet(df_august)

future_august <- make_future_dataframe(m_august, periods = 1, freq = "month")

forecast_august <- predict(m_august, future_august)

august_2017_forecast <- forecast_august$yhat[length(forecast_august$yhat)]

print(july_2017_forecast)
print(august_2017_forecast)

modelTester(july_2017_forecast, august_2017_forecast)
```

```{r}
library(forecast)
library(dplyr)
library(lubridate)

july_df <- myHotel_bookings_completed_Jul17 %>%
  filter(hotel == "Resort Hotel", is_LMC == 1) %>%
  group_by(year = year(arrival_date), month = month(arrival_date)) %>%
  summarise(LMCs = sum(is_LMC)) %>%
  arrange(year, month)

july_LMCs_ts <- ts(july_df$LMCs, start=c(july_df$year[1], july_df$month[1]), frequency=12)

model_july <- auto.arima(july_LMCs_ts)

forecast_july <- forecast(model_july, h=1)

forecast_value_july <- round(forecast_july$mean)

august_df <- myHotel_bookings_completed_Aug17 %>%
  filter(hotel == "Resort Hotel", is_LMC == 1) %>%
  group_by(year = year(arrival_date), month = month(arrival_date)) %>%
  summarise(LMCs = sum(is_LMC), .groups = 'drop') %>%
  arrange(year, month)

august_LMCs_ts <- ts(august_df$LMCs, start=c(august_df$year[1], august_df$month[1]), frequency=12)

model_august <- auto.arima(august_LMCs_ts)

forecast_august <- forecast(model_august, h=1)

forecast_value_august <- round(forecast_august$mean)
july_forecast_value <- round(forecast(model_july, h=1)$mean)
august_forecast_value <- round(forecast(model_august, h=1)$mean)

numeric_forecast_value_july <- round(forecast_july$mean[1])
numeric_forecast_value_august <- round(forecast_august$mean[1])

print(numeric_forecast_value_july)
print(numeric_forecast_value_august)
modelTester(numeric_forecast_value_july, numeric_forecast_value_august)
```

# Reasoning

Prophet: The script filters the hotel bookings data for "Resort Hotel" and then groups the data by year and month to calculate the sum of LMCs. It creates a dataframe suitable for the Prophet model, with ds representing the date (year-month) and y representing the number of LMCs. I chose this model due to it's simplicity and effictiveness in predicting prowess, though I am not fully sure if it was perfectly applicable in this instance it was much simpler to set up after various reviewing on documentation.

Arima: Similar to the Prophet method, this approach also focuses on filtering and summarizing LMCs for "Resort Hotel." However, it explicitly uses the year and month functions from the lubridate package to extract these units from the arrival date. This method directly ties the time series object to specific time units, facilitating the ARIMA model's understanding of the data's temporal structure. The auto.arima function performs an automated search over possible model specifications, this is the model that I thought sounded better but was more complex in putting together.
